---
title: "Practical Machine Learning Assignment"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Summary
People regularly quantify how much activity they do, but they rarely quantify how well they do it. In this project, the goal is to analyze data from accelerometers on the belt, forearm, arm, and dumbell of six participants. They were asked to perform barbell lifts correctly and incorrectly in five different ways. ar

The goal of this machine learning exercise is to predict the manner in which the participants did the exercise, that is to predict the "classe" variable found in the training dataset. The prediction model will then  used to predict 20 different test cases.


##Data Processing and Analysis
Loading the required libraries and reading in the training and testing datasets, assign missing values to values that are 'NA' or blank.
```{r}
set.seed(1)
```

```{r echo=FALSE}
setwd("C:/myFile/coursera/m8/project")

```

```{r}

datTraining<-read.csv("pml-training.csv", header=TRUE, na.strings=c("NA", ""))
datTesting<-read.csv("pml-testing.csv", header=TRUE, na.string=c("NA", ""))

```

##Data Cleaning
Excluding data with "NA" and delete additional unnecessary columns 

```{r}
library(caret)
        
datTrainNoNA<-datTraining[, apply(datTraining, 2, function(x) !any(is.na(x)))] 
dim(datTrainNoNA)

datTrainingClean<-datTrainNoNA[,-c(1:8)]
dim(datTrainingClean)

datTrainNzv <- nearZeroVar(datTrainingClean, saveMetrics=TRUE)
datTrainingCleanNew <- datTrainingClean[,datTrainNzv$nzv==FALSE]
dim(datTrainingCleanNew)
barplot(table(datTrainingCleanNew$classe),ylab="count",xlab="classes",col="green",main="count by classes")
```

##Data Partitioning and Prediction Process
Split the clean training dataset into a training dataset (70% of the observations) and a validation dataset (30% of the observations). This validation dataset is use to perform cross validation when developing the model.

```{r}
inTrainingData<-createDataPartition(y=datTrainingCleanNew$classe, p=0.70,list=FALSE)
myTrainData <- datTrainingCleanNew[inTrainingData,]
dim(myTrainData)

myTestData <- datTrainingClean[-inTrainingData,] 
dim(myTestData)
```

##Prediction Model
Prediction with Random Forests
```{r}

library(randomForest)

fitControl <- trainControl(method="cv", number=5, allowParallel=T, verbose=T)
rfModel<-train(classe~.,data=myTrainData, method="rf", trControl=fitControl, verbose=F)
print(rfModel)

```

#Validate the Model
Using confusionMatrix to evaluate and cross validate the model
```{r}
predrf<-predict(rfModel, myTestData)
confusionMatrix(predrf, myTestData$classe)
```


# Result of the prediction to test dataset
The accuracy is 99%, thus the predicted accuracy for the out of sample error is 1%. Random Forests is chosen
because it produce better results. 
```{r}
predictpmlTesting<-predict(rfModel, datTesting)

#Output for the prediction of the 20 cases provided
predictpmlTesting
```
20/20 cases from the quiz were correctly predicted



